---
title: "Explore NEON LiDAR Data"
author: "Kristin Braziunas"
date: "June 15, 2016"
output: html_document
---

# About My Project

### Bio

I am an incoming MS/PhD Student (starting this fall) studying landscape ecology at the 
University of Wisconsin–Madison (Dept of Zoology), where I started working in January 2016 
as a Research Specialist in Forest Simulation Modeling. Over the course of my graduate 
studies, I hope to contribute to research on the impacts of changing climate and fire 
regimes on forest ecosystems, as well as how these impacts might change the way we 
approach forest and fire management. I am particularly interested in using multiple 
computational models and quantitative analyses to simulate forest development under novel 
environmental conditions at different spatial and temporal scales. Currently, I am 
parameterizing and testing the process-based model iLand for subalpine conifer forests in 
Yellowstone.

### Goals

* Learn more about the data that NEON will collect and disseminate
* Use remote sensing data to characterize forest structure and composition
* Refine existing and develop new skills in coding and data analysis

### Capstone Project

**Question: How well does remote sensing data represent forest structure and composition 
for initialization of a forest landscape simulation model?**

Distance-dependent, individual-tree, process-based forest landscape simulation models are 
needed to simulate potential forest growth and structure trajectories in a future 
characterized by novel environmental drivers (e.g. changing climate, disturbance frequency 
and severity). These models require detailed initialization files including information on 
forest structure, species composition, and individual tree characteristics. Collecting 
field data for model initialization is labor intensive at the plot scale, and unrealistic 
at the landscape scale. However, remote sensing of forest structure and species 
composition can potentially provide accurate and complete data at the landscape scale. I 
will first derive forest structure and species composition from remote sensing data and
compare it to stands in the same geographic region from the Forest Inventory and Analysis
(FIA) Program. For example, I could look at dominant tree heights, canopy cover, and 
percent species composition.

Time permitting, I will then derive individual tree characteristics (e.g. DBH, height)  
from remote sensing data for comparison with FIA stands. The long-term goal of this 
exercise is to use remote sensing data to create initialization files for 
the process-based model iLand, which requires the following information on individual 
trees and saplings: species, density, DBH, height, and age. When available, soil, climate, 
and atmospheric data (e.g. nutrient content in soils, CO2 concentration) will also be used 
for model initialization. 

Data needed:
* LiDAR (vegetation structure, heterogeneity, topography) 
* Hyperspectral (vegetation cover, species composition) 
* FIA data (condition, plot, and tree tables)

Data availability: Remote sensing data is available for sites in D01 (Northeast) and D03 
(Southeast). FIA data is available for all states via the FIA DataMart except Hawaii.

***

# Week 3 Activity

For this activity, I am exploring NEON LiDAR data available for Lower Teakettle, a 
relocatable terrestrial field site in NEON Domain 17 (Pacific Southwest). [Learn more
about the Lower Teakettle field
site.](http://www.neonscience.org/science-design/field-sites/lower-teakettle)

I will be specifically looking at the Digital Surface Model (DSM) output on site elevation and the Canopy Height Model (CHM) output on vegetation heights.

First, I'm going to set up my working directory.

```{r setup-read-data, message=FALSE}
library(rgdal)
library(raster)

# set working directory
setwd("~/Documents/Data/NEONDI-2016/")

```

And then load the Teakettle Digital Surface Model data.

```{r load-dsm-raster}

# import dsm
teak_dsm <- raster("NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarDSM.tif")

```

Initial plot of Teakettle DSM.

```{r plot-dsm-raster, echo=FALSE}

# plot dsm
plot(teak_dsm, main=" Digital Surface Model (DSM) \n NEON Lower Teakettle Site ",
     col=rainbow(10))

```

Now, I'll import the Teakettle Canopy Height Model raster.

```{r load-chm-raster}

# import chm
teak_chm <- raster("NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

```

Initial plot of Teakettle CHM.

```{r plot-chm-raster}

# plot chm
plot(teak_chm, main=" Canopy Height Model (CHM) \n NEON Lower Teakettle Site ")

```

From the DSM plot and histogram, it appears that this site is mostly sloped with a taller peak near the northwest corner of the site and also a point of slightly higher elevation in the southeast. From the CHM plot, it appears that the site does not have a lot of vegetation.

Plot histograms of Teakettle DSM and CHM.

```{r plot-hist-dsm-chm, warning=FALSE}

# histogram of dsm
hist(teak_dsm, main=" DSM Elevation Frequency \n NEON Lower Teakettle Site ",
     breaks=10, col="darkkhaki", xlim=c(2150,2400), ylim=c(0,30000),
     xlab="Elevation (m)")

# histogram of chm
hist(teak_chm, main=" CHM Height Frequency \n NEON Lower Teakettle Site ", 
     col="forestgreen", axes=FALSE, xlab="Height (m)", ylab="Frequency (thousands)",
     xlim=c(0,60))

# change y axis to thousands, keep x axis same
axis(2, at=axTicks(2), labels=paste(axTicks(2)/1000))
axis(1, at=axTicks(1), labels=paste(axTicks(1)))

```

The CHM histogram is not very interesting - it looks like it is only capturing the frequency when height equals 0. I'll try modifying the CHM raster to eliminate values equal to 0 to see if I can plot the frequencies of vegetation heights that are present.

```{r modify-chm-hist, warning=FALSE}

# modify chm raster so only values >0
summary(teak_chm)
teak_chm.mod <- teak_chm
teak_chm.mod[teak_chm.mod == 0] <- NA
summary(teak_chm.mod)

# histogram of modified chm
hist(teak_chm.mod, main=" CHM Height Frequency \n NEON Lower Teakettle Site ", breaks=5,
     col="forestgreen", axes=FALSE, xlab="Height (m)", ylab="Frequency",
     xlim=c(0,60), ylim=c(0,100))

# change y axis to thousands, keep x axis same
axis(2, at=axTicks(2), labels=paste(axTicks(2)))
axis(1, at=axTicks(1), labels=paste(axTicks(1)))

```

From the summary and the histogram, it looks like the vegetation that is present only
within 50-56m in height, with a median height of 51.14m.