---
title: "Explore NEON LiDAR Data"
author: "sjmc"
date: "June 9, 2016"
output: 
  html_document: 
    theme: spacelab
---

### Short Bio
Hi, I'm a Ph.D. student and participant at the 2016 NEON Data Institute. <span style="color:green">__Forest ecology__</span> is my jam. Much of my research looks at how organisms and ecosystems respond to stress, such as drought and air pollution. I'm also interested in what goes on in ecosystems after disturbances. I work at the University of California Berkeley, and my field studies are in the Sierra Nevada. My full bio is available [here.](https://en.wikipedia.org/wiki/Triceratops)
	
### About My Workshop Project
  * I'm interested in investigating how a disturbance - detected with hyperspectral veg data or simply known from ground measurements - is reflected in CO2 or other fluxes.
    * Hoping to work w/ data from the Pacific Southwest or Pacific Northwest domains, and am especially interested in Teakettle, though I know it is still getting up and running. 
    * Data needed may include:
        - hyperspectral vegetation data
        - standardized forest structure measurements (DBH, height, maybe biomass)
        - fluxes for pre- and post- event
        - (External) disturbance event information, such as a fire or mortality event.
    * In the process, I'm also interested in quantifying uncertainties when combining vegetation plots or LiDAR w/ remotely sensed data. 
  
### Week 3 Activity
  * This page is a demo of a few NEON LiDAR products using Teakettle Experimental Forest as an example
    - And, practice using __RMarkdown__

#####1. First, let's get set up:
```{r setup-read-data, message=FALSE }
  #load libraries
    library(rgdal)
    library(raster)
   # set working directory to ensure R can find the file we wish to import
   setwd("/Users/stellac/Documents/HomeWork/data/NEONDI-2016/")
```
* Note that in order to knit properly, your .md file (this document) should be saved in that same wd. At least, that seems to be what the knittr documentation says. 

#####2. Load the Teakettle elevation raster:
```{r load-dsm-raster }
   # import dsm
    teak_dsm <- raster("NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarDSM.tif" )
   
```
#####3. View the Teakettle elevation raster (digital surface model):
```{r plot-dsm-raster }

   # plot it up flat
   plot(teak_dsm, main= "Teakettle Digital Surface Model (m)")
   
```

+ Observations: looks like a ridgetop in the upper left. At this elevation, could be bare granite.
+ High points are probably trees.

#####4. View that surface model as a histogram:
```{r histo-dsm-raster, warning=FALSE}
   # plot it up histogram
   hist(teak_dsm,breaks=20,col="pink",ylim = c(0,20000),main = "Teakettle Elevation Histogram (m)",xlab = "Elevation (meters)",ylab="Frequency")

```

#####5. Load the Teakettle LiDAR-derived canopy height model:
```{r load-CHM-raster }
   # import CHM
   teak_CHM <- raster("NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif" )
```
#####6. View the Teakettle LiDAR-derived canopy height model:
```{r plot-CHM-raster }
   # plot it up flat
#note to self : tried it w/ this added, but still just a few specks
#breaks = c(seq(0,70,by=5)), col=terrain.colors(14))
   plot(teak_CHM, main= "Teakettle Canopy Height Model")
```

+ Observations: seems like this might be a location w/ very few mature trees? Not sure how to interpret this, I'd expect a dense forest - especially based on the DSM. Not a very large area, though.

#####7. Canopy height model as a histogram, log-log scale
```{r histo-CHM-histo, warning=FALSE }
   # plot it up histogram-wise
   #log scale to see detail/variation 
   hist(log(teak_CHM), breaks=50,col="yellow",main = "Teakettle Canopy Height Model Histogram",ylab = "Log Frequency", xlab = "log height (m)")
```

 + that's all for now!
