---
title: "Week 3 assignment"
output: html_document
---

## Max Joseph

I am part of the Analytics Hub with Earth Lab and Research Computing at the University of Colorado Boulder.
My background is in hierarchical Bayesian modeling and disease dynamics, but much of my current work involves writing scientific software to make data more useful.
I am hoping to become more proficient with remotely sensed data by taking this course, so that I can more effectively integrate data from different sources.

## Project summary

Did fire history influence erosion in the 2013 Boulder floods? 
Because fire destroys soil-stabilizing vegetation, it is possible that *all else being equal*, areas that had recently burned prior to the 2013 floods may have experienced more erosion.

I propose to investigate the relationship between burn history and erosion using the 2013 pre and post flood NEON LiDAR data. 
Boulder County provides shapefiles of historical fire boundaries, with information on fire age/time since ignition for 33 fires, spanning the years 1978 to 2012.
We could investigate binning the landscape so that comparisons among areas are valid, for instance considering only regions with comparable aspect, slope, and hydrogeographic traits. 
Eventually, we would like to compare the distributions of pixel-wise elevation changes for each area, mapping the distribution in some way to fire age (e.g., a plot with fire age on the x-axis and mean absolute change in elevation on the y-axis). 

Alternatively, if the data are available, an even better approach would involve extracting vegetation data from the LiDAR data. 
For instance, we specifically expect that large live trees would be particularly stabilizing. 
If we could link the height of plants to the amount of elevation change due to erosion, all using the LiDAR data, then probably we would have a better resolved understanding of the soil-stabilizing effects of vegetation. 

## Loading lidar rasters

Let's see what files are available in the lidar folder.

```{r list_files}
library(raster)
library(rasterVis)

# what data are available?
data_dir <- "~/Downloads/NEONDI2016TEAK/NEONdata/D17-California/TEAK/2013/lidar/"
files <- list.files(data_dir)
files
```

There are five tif files in the lidar folder. Let's load them as a list.

```{r load_files}
paths <- paste0(data_dir, files)
r_list <- lapply(paths, raster)
r_list
```

## Creating a raster stack

Each raster in the list has the same projection, extent, and resolution.
We can stack these rasters to make a RasterStack. 

```{r make_stacks}
r_stack <- stack(r_list)
r_stack
```

## Plotting the rasters

To plot these, I'll use some functions from RasterVis.

```{r plot_stack}
plot(r_stack)
```

How correlated are the values among the layers?

```{r splom_stack}
splom(r_stack)
```

How about a level plot?

```{r terrible_levelplot}
levelplot(r_stack)
```

Looks horrible, because all of the values among rasters are on different scales. 

```{r bad_levelplot}
levelplot(scale(r_stack))
```

Less horrible, but still bad because of outliers in the CHM (crown height) layer, which has a long right tail (a few pixels with very large crown height values). 

## Generating raster histograms

We can use the `hist` function, but I'll suppress warnings.

```{r histograms, warning=FALSE}
hist(r_stack)
```

## Discussion

The `raster` package makes all of these operations fairly easy as long as everything has the same extent, projection, and resolution. 
R markdown is nice because the documents are portable. 
